Journal and article coverage

### 2025-03-18 analysis/manuscript.qmd
Some minor language fixes

```{r dependencies}
library(tidyverse, warn.conflicts = FALSE)
library(bigrquery)
library(kableExtra)
library(patchwork)
library(ComplexUpset)
library(ggrepel)
library(ggpp)

@@

This study investigates the suitability of open scholarly data sources for assessing the impact of transformative agreements on hybrid open access.
As shown in Figure \ref{fig:workflow}, the methodology involved comparing hoaddata, an openly available collection of open research information on hybrid open access, with the bibliometric databases Web of Science and Scopus.
This section introduces the initial data sources, followed by a presentation of the necessary data processing steps to obtain eligible articles from transformative agreements by using author roles (first and corresponding) and harmonised affiliation data.
<!-- Overall, xxxx hybrid journals from xxx agreements that published at least one open access article between 2019 and 2023. formed the basis of this study. -->

@@

hoaddata, developed and maintained by the author to support open access monitoring and research [@Jahn_2025], is an R data package collecting information on hybrid open access uptake from multiple openly available data sources.
It combines article-level metadata from Crossref and OpenAlex with transformative agreement information from the cOAlition S Journal Checker Tool (JCT), which links journal and institutional data to agreements in the ESAC registry.

As discussed elsewhere [@Jahn_2025], hoaddata uses a unified dataset of historic JCT snapshots, which were safeguarded since July 2021.
Journal data were enriched with ISSN variants linked to an ISSN-L and insitution data with associated ROR-IDs representing university hospitals or institutes of large research organisations, according to OpenAlex' institution entity.
A comprehensive exclusion of fully open access journals using multiple journal lists (DOAJ, OpenAlex, Bielefeld)  was performed to identify hybrid journals.
hoaddata relies on Crossref for obtaining journal publication volume and open access status through Creative Commons licence information relative to the published version ("version of record").
Because of limited affiliation metadata in Crossref [@van_Eck_2022], hoaddata sources  affiliation metadata from OpenAlex.

@@

hoaddata follows good practices for computational reproducibility using R [@Marwick_2018].
The package, which includes data, code, a test suite and documentation, is openly available on GitHub.
To ensure computational reproducibility while aggregating the data, a GitHub Actions continuous integration and delivery (CI/CD) workflow handles data retrieval from the SUB Göttingen's open scholarly data warehouse based on Google BigQuery, which provides high-performant programmatic access to monthly snapshots of Crossref and OpenAlex.
The workflow has run regularly to fetch updates from these data sources since 2022.
The package version used in this study is 0.3, containing data from the Crossref 2024-08 dump provided to Crossref Metadata Plus subscribers and the OpenAlex 2024-08-29 monthly dump.
This version including the computation log is available on GitHub (https://github.com/subugoe/hoaddata/releases/tag/v.0.3).

@@

Clarivate Analytics' Web of Science (WoS) is a well-established proprietary bibliometric database consisting of several collections [@Birkle_2020]. 
The collections considered in this study were the Science Citation Index Expanded (SCIE), the Social Sciences Citation Index (SSCI) and the Arts & Humanities Citation Index (AHCI), collectively referred to as Web of Science Primary (WoS Primary).

The WoS Primary provides important data points for analysing open access: author affiliations and roles, differentiation of journal articles into document types representing different types of journal contributions, such as original articles or reviews, and open access status information derived from OurResearch's Unpaywall, the same provider as OpenAlex.
However, it lacks information about journals and articles under transformative agreements.

For programmatic access to article-level data, this study used the database of the Kompetenznetzwerk Bibliometrie (KB) in Germany. The KB processes raw XML data provided by Clarivate Analytics, which is provided as an in-house PostgreSQL database under a uniform schema.
To support reproducibility, KB maintains annual snapshots of the database. 
Accordingly, this study used the annual snapshot from April 2024 (wos_b_202404), which is considered to cover almost the entire previous publication year [@schmidt_2024_13935407].

@@

The hoaddata package (v.0.3) provided an enriched and unified list of hybrid journals in transformative agreements according to the JCT, covering agreements active from July 2021 to July 2024 according to the ESAC registry.
<!-- This study examined the adoption of open access for hybrid journals included in transformative agreements, which published at least one journal article in the five-year period 2019-2023. 
The initial data source was the Transformative Agreements Public Data from the cOAlition S Journal Checker Tool (JCT), which helps authors identify suitable open access publishing venues, as provided by hoaddata.
 Weekly snapshots of this data were archived using an automated GitHub Actions workflow (https://github.com/njahn82/jct_data) and into hoaddata.
The JCT data used in the examined hoaddata version, which was enriched with linking ISSN (ISSN-L) and ROR-IDs from associated institutions before inclusion, cover agreements active from July 2021 to July 2024. -->

<!-- The JCT journal data was enriched with the linking ISSN (ISSN-L) according to the ISSN Registry Agency.
A comprehensive exclusion of fully open access journals was performed, following the methodology described in @jahn_2025.
This involved checking multiple journal lists (DOAJ, OpenAlex, Bielefeld) and article-level investigations.
The JCT institution data was enriched with ROR-IDs from associated institutions, such as university hospitals or institutes of large research organisations such as the Max Planck Society, according to OpenAlex' institution entity.
<!-- In total, xxxx hybrid journals could be linked to xxx agreements listed in the ESAC Transformative Agreements Registry, which formed to basis for obtaining article-level data including author affiliations, open access status ans document types. -->
Article metadata were retrieved from each database using all ISSN variants linked to an ISSN-L. The metadata included DOIs, publication dates, article types, open access information as well as author roles and institutional affiliations.
Publication years were determined using the earliest known date of publication in a journal.
In hoaddata, this corresponded to Crossref's issued date.<!--, although variations were observed (check ref)-->
For Web of Science and Scopus, the earliest publication date was used where available, with Scopus dates specifically determined by the KB through version tracking of the raw data.

Many transformative agreements typically cover only certain types of journal
articles, in particular original research articles including reviews.
Because of limited information on these document types in open scholarly data [@Haupka_2024], hoaddata used an extended version of Unpaywall's paratext recognition approach to exclude non-scholarly content [@Piwowar_2018].
To exclude conference supplements, which are also often not covered by transformative agreements, only articles published in regular issues, indicated by numerical pagination, were considered.
For Web of Science and Scopus, their established document type classifications were used to identify original research articles and reviews, referred to as core articles throughout this study.

@@

Published articles were considered hybrid open access  when they were made freely available under a Creative Commons license on publishers' platforms. 
While hoaddata sourced this information from Crossref license metadata, Web of Science and Scopus relied on Unpaywall. 
Unpaywall supplements Crossref metadata by parsing publisher websites directly, addressing cases where publishers do not provide machine-readable Creative Commons license information [@Piwowar_2018]. 
This additional parsing remains necessary despite transformative agreement workflows recommend the depositing license information during DOI registration [@Geschuhn_2017].
Both Web of Science and Scopus defined hybrid open access consistently as content available under Creative Commons licenses on publisher platforms, distinguishing it from bronze open access that lack such explicit license information.

@@

```{r cache=TRUE}
# Article level data TA
ta_oa_inst <- readr::read_csv(here::here("data-raw", "ta_oa_inst.csv"))
```
Author affiliations were retrieved for both first and, if available, corresponding authors across all databases to prepare the linking between articles and institutions covered by transformative agreements.
To handle different address variants, database-specific affiliation identifiers were used: ROR-IDs from OpenAlex for hoaddata, affiliation enhanced names for Web of Science, and Scopus Affiliation Identifiers for Scopus.
Additionally, ISO country codes were retrieved for each author's address to compile country-level statistics. 
These country codes for Web of Science and Scopus were provided by the KB.

Because neither Web of Science nor Scopus support ROR IDs—the institution identifier used by the JCT—a two-step matching process was implemented to harmonise affiliation data.
First, `r length(unique(ta_oa_inst$doi))` articles from `r length(unique(ta_oa_inst$ror))` institutions with ROR-IDs in the JCT data since 2017 (according to hoaddata) were processed to map first authors' ROR-IDs to corresponding proprietary affiliation identifier in Web of Science and Scopus using DOI matching.
Then, an algorithm selected the most frequent ROR ID and proprietary identifier pairs to handle multiple affiliations and organisational hierarchy differences.

@@

As a result of the comprehensive data processing described above, datasets on open access in hybrid journals included in transformative agreements were aggregated for each database at country and journal level by year.
@tbl-methods_overview_tab provides a general overview of the coverage between 2019 and 2023 per database.
It shows that the majority of hybrid journals published at least one original research article or review marked as core during the five-year period.
These journals formed the basis for the subsequent calculation of article-level indicators.

```{r methods_overview_tab}
#| label: tbl-methods_overview_tab
#| tbl-cap: "Coverage of hybrid journals in transformative agreements 2019-23."
load(here::here("data", "manuscript_data.Rda"))
methods_overview_table_df <- bind_rows(
  # Journals
  jns_ind,
  # Articles
  articles_ind,
  # OA
  oa_articles_ind,
  # Affiliation and Auhtor roles
  aff_ind
) |>
  # Human-readable indicator labels
  mutate(
    indicator = case_when(
      indicator == "ind_active_jns" ~ "Active journals",
      indicator == "ind_active_jns_core" ~ "Active journals (core)",
      indicator == "ind_active_jns_core_with_oa" ~ "Active journals (core) with OA",
      indicator == "ind_articles_total" ~ "Total published articles ",
      indicator == "ind_articles_core" ~ "Core articles",
      indicator == "ind_articles_doi" ~ "Articles with DOI",
      indicator == "ind_articles_core_doi" ~ "Core articles with DOI",
      indicator == "ind_oa_articles_all" ~ "OA articles",
      indicator == "ind_oa_articles_core" ~ "Core OA articles",
      indicator == "first_aff" ~ "First author articles",
      indicator == "cor_aff" ~ "Corresponding author articles",
    )
  )

kableExtra::kbl(
  methods_overview_table_df,
  booktabs = TRUE, 
  format.args = list(big.mark = ","), 
  digits = 1, 
  escape = TRUE, 
  position = "H",
  col.names = c("", "HOAD", "Web of Science", "Scopus")
) |>
  kable_styling(position = "center") |>
  row_spec(0, bold = TRUE) |>
  pack_rows("Hybrid journal coverage", 1, 3) |>
  pack_rows("Publication volume", 4, 5) |>
  pack_rows("Digital Object Identifier (DOI) coverage", 6, 7) |>
  pack_rows("Open Access (OA) metrics", 8, 9) |>
  pack_rows("Core articles with affiliation data", 10, 11) 
```

While hoaddata only covers articles with a DOI, Scopus and Web of Science publication indicators were calculated using their database identifier.
A subsequent comparison of DOI coverage shows that non-core articles in Web of Science often lacked a DOI.
This was particularly the case for meeting abstracts, which are notably prevalent in Health Sciences journals [@Melero_Fuentes_2025] and are not indexed by Scopus [@Donner_2017].
Open access indicators were aggregated by DOI, as Unpaywall only collects information on open access status for articles with a DOI.
A closer look at the core articles with affiliation data reveals a lack of data on corresponding authors in the case of OpenAlex, the affiliation data source used by hoaddata, compared to Web of Science and Scopus.
Only about two-thirds of the articles examined provided corresponding author affiliations.
For first authors, the proportion was `r methods_overview_table_df |> filter(indicator == "First author articles") |> pull(hoad) / methods_overview_table_df |> filter(indicator == "Core articles with DOI") |> pull(hoad) * 100`%.
Therefore, only first author data for hoaddata were considered in the following analysis.

@@

upset_overview <- ComplexUpset::upset(
  upset_data,
  c("Crossref", "WoS", "Scopus"),
  # Only journals in Crossref
  min_size = 30,
  set_sizes = FALSE,
  stripes = "transparent",
  # wrap = TRUE,
  name = "Data Source Coverage",
  base_annotations = list(
    "Journals" = (
      ComplexUpset::intersection_size(
        counts = FALSE,
        text = list(size = 8 / .pt),
        text_mapping = aes(
          label = paste0(!!upset_text_percentage(), '\n(', format(!!size, big.mark = ","), ')'),
          colour = ifelse(!!size > 5000, 'on_bar', 'on_background'),
          y = ifelse(!!size > 5000, !!size - 4000, !!size)
        ), width = bar_width
      ) +
        scale_y_continuous(labels = scales::number_format(big.mark = ",")) +
        ggtitle("A")
    )
  ),
  annotations = list(
    "Journal size\n(log scale)" = (
      ggplot(mapping = aes(x = intersection, y = overlap)) +
        geom_boxplot(outliers = FALSE, width = bar_width) +
        scale_y_log10(labels =  scales::number_format(big.mark = ","), limits = c(5, 10000)) +
        ggtitle("C")
    ),
    "Articles" = (
      ggplot(mapping = aes(y = n_cr)) +
        geom_bar(
          stat = "identity",
          color = "grey80",
          width = bar_width,
          aes(fill = "Surplus")
        ) +
        geom_bar(
          aes(y = overlap, fill = "Shared"),
          stat = "identity",
          width = bar_width,
          color = "#0091c8"
        ) +
        scale_fill_manual(
          values = c(
            `Surplus` = "grey80",
            `Shared` = "#0091c8"
          ),
          name = "DOI"
        ) +
        scale_y_continuous(labels =  scales::number_format(big.mark = ","),
                           breaks = c(0, 2500000, 5000000, 7500000)) +
        guides(fill = guide_legend(reverse = TRUE)) +
        ggtitle("B")
    )
  ),
  matrix = (intersection_matrix(geom = geom_point(
    shape = "square", size = 2
  )))
)
upset_overview & my_upset_theme
```
Journal coverage analysis revealed that `r round(nrow(union_jns) / nrow(all_jns) * 100)`% (n = `r nrow(union_jns)`) of hybrid journals included in transformative agreements were indexed in all three databases (Figure \ref{fig-upset_coverage_results}A).
The second-largest set consisted of journals indexed in both hoaddata and Scopus, comprising `r round( jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |>  nrow()`) of hybrid journals.
Notably, `r round(jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |>  nrow()`) of journals were exclusively contained in hoaadata, while `r round( jn_upset_articles |> filter(Crossref == FALSE, Scopus == TRUE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(Crossref == FALSE, Scopus == TRUE) |>  nrow()`) were only found in the proprietory database Scopus. 
Of these, `r jn_upset_articles |> filter(WoS == TRUE, Crossref == FALSE, Scopus == TRUE) |>  nrow()` were also available in the Web of Science.
Upon inspection, this group of hybrid journals included only in the proprietary data sources mainly represents hybrid journals for which no open access evidence could be retrieved from Crossref, which served as the open access evidence source for hoaddata.

In terms of article coverage, Figure \ref{fig-upset_coverage_results}B shows the total publication volume per combination in terms of DOI availability.
The largest set of hybrid journals, which includes all three data sources, also contains the largest number of articles. 
In total, these journals recorded `r upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum()` overlapping articles, represented by the blue bar. They represent `r round(upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum() / methods_overview_table_df |> filter(indicator == "Core articles with DOI") |> pull(wos) * 100)`% of core articles with DOI indexed in the Web of Science sample, and `r round(upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum() / methods_overview_table_df |> filter(indicator == "Core articles with DOI") |> pull(scp) * 100)`% in Scopus.
Another `r upset_data |> filter(Crossref == TRUE, WoS == FALSE, Scopus == TRUE) |> pull(n_cr) |> sum()` articles were exclusive to both Scopus and hoaddata. Exclusively in hoaddata were `r upset_data |> filter(Crossref == TRUE, WoS == FALSE, Scopus == FALSE) |> pull(n_cr) |> sum()` articles, and exclusively in the proprietary databases were `r upset_data |> filter(Crossref == FALSE, Scopus == TRUE) |> pull(overlap) |> sum(na.rm = TRUE)` articles.

@@

Figure \ref{fig-upset_coverage_results}B also shows DOIs that were only available via hoaddata (grey area).
In case of hybrid journals covered by all three data sources, `r miss_dois |> nrow()` DOIs were only present in hoaddata.
After validation at the DOI level using the KB databases and manual inspection, the main reasons for missing DOI coverage in the proprietary database were insufficient classification of journal content as original research articles and reviews during the compilation of hoaddata. 
Particulalry, letters and editorials could not be fully detected.
Paratext recognition failed in `r round(miss_dois_val |> filter(is.na(scp_item_type), is.na(wos_item_type)) |> pull(n_doi) / miss_dois |> nrow() * 100)`% of DOIs to identify non-scientific content such as front matters or reviewer lists, which are generally not indexed by Scopus and Web of Science. 
To a lesser extent, differences in publication and indexing dates were a reason for non-overlapping DOIs.

Using overlapping DOIs, the publication volume between 2019 and 2023 was also calculated for each journal. 
Figure \ref{fig-upset_coverage_results}C illustrates the distribution for each combination. It shows that there is a large spread across the journals covered by all three data sources. 
Furthermore, journals in this set published more on average than journals covered by only one or two data sources. 
In particular, the journals covered exclusively by hoaddata are substantially smaller than those covered by all three sources. 
Upon inspection, these are often newly launched journals, which explains the relatively low five-year publication volume.

@@

```{r oa_country_comparision_plot_, fig.asp=1, fig.width = 8}
layout <- "EE
AC
BD"

>>>>

```{r dependencies}
library(tidyverse, warn.conflicts = FALSE)
library(bigrquery)
library(kableExtra)
library(patchwork)
library(ComplexUpset)
library(ggrepel)

@@

This study investigates the suitability of open scholarly data sources for assessing the impact of transformative agreements on hybrid open access.
As shown in Figure \ref{fig:workflow}, the methodology involved comparing hoaddata, an openly available collection of open research information on hybrid open access, with the bibliometric databases Web of Science and Scopus.
This section introduces the initial data sources, followed by a presentation of the necessary data processing steps to obtain eligible articles enabled by transformative agreements using author roles (first and corresponding) and harmonised affiliation data.
<!-- Overall, xxxx hybrid journals from xxx agreements that published at least one open access article between 2019 and 2023. formed the basis of this study. -->

@@

hoaddata, developed and maintained by the author to support open access monitoring and research [@Jahn_2025], is an R data package collecting information on hybrid open access uptake from multiple openly available data sources [@hoaddata].
It combines article-level metadata from Crossref and OpenAlex with transformative agreement information from the cOAlition S Journal Checker Tool (JCT), which links journal and institutional data to agreements in the ESAC registry.

As discussed elsewhere [@Jahn_2025], hoaddata uses a unified dataset of various safeguarded JCT snapshots.
Journal data were enriched with ISSN variants linked to an ISSN-L and institution data with associated ROR-IDs representing university hospitals or institutes of large research organisations, according to OpenAlex' institution entity.
A comprehensive exclusion of fully open access journals using multiple journal lists (DOAJ, OpenAlex, Bielefeld)  was performed to identify hybrid journals.
hoaddata relies on Crossref for obtaining journal publication volume and open access status through Creative Commons licence information relative to the published version ("version of record").
Because of limited affiliation metadata in Crossref [@van_Eck_2022], hoaddata sources  affiliation metadata from OpenAlex.

@@

hoaddata follows good practices for computational reproducibility using R [@Marwick_2018].
The package, which includes data, code, a test suite and documentation, is openly available on GitHub.
To ensure computational reproducibility while aggregating the data, a GitHub Actions continuous integration and delivery (CI/CD) workflow handles data retrieval from the SUB Göttingen's open scholarly data warehouse based on Google BigQuery, which provides high-performant programmatic access to monthly snapshots of Crossref and OpenAlex.[^subugoe]
The workflow has run regularly to fetch updates from these data sources since 2022.
The package version used in this study is 0.3, containing data from the Crossref 2024-08 dump provided to Crossref Metadata Plus subscribers and the OpenAlex 2024-08-29 monthly dump.
It covers agreements collected between July 2021 to July 2024 from the JCT.
This version including the computation log is available on GitHub (https://github.com/subugoe/hoaddata/releases/tag/v.0.3).

@@

Clarivate Analytics' Web of Science (WoS) is a well-established proprietary bibliometric database consisting of several collections [@Birkle_2020]. 
The collections considered in this study were the Science Citation Index Expanded (SCIE), the Social Sciences Citation Index (SSCI) and the Arts & Humanities Citation Index (AHCI).

These collections provides important data points for analysing open access: author affiliations and roles, differentiation of journal articles into document types representing different types of journal contributions, such as original articles or reviews, and open access status information derived from OurResearch's Unpaywall, the same provider as OpenAlex.
However, the Web of Science lacks information about journals and articles under transformative agreements.

For programmatic access to article-level data, this study used the database of the Kompetenznetzwerk Bibliometrie (KB) in Germany. The KB processes raw XML data provided by Clarivate Analytics, which is ingested into an in-house PostgreSQL database under a uniform schema.
To support reproducibility, KB maintains annual snapshots of the database. 
Accordingly, this study used the annual snapshot from April 2024 (wos_b_202404), which is considered to cover almost the entire previous publication year [@schmidt_2024_13935407].

@@

Using an enriched and unified list of hybrid journals in transformative agreements according to the JCT, article metadata were retrieved from each database using all ISSN variants linked to an ISSN-L.
<!-- This study examined the adoption of open access for hybrid journals included in transformative agreements, which published at least one journal article in the five-year period 2019-2023. 
The initial data source was the Transformative Agreements Public Data from the cOAlition S Journal Checker Tool (JCT), which helps authors identify suitable open access publishing venues, as provided by hoaddata.
 Weekly snapshots of this data were archived using an automated GitHub Actions workflow (https://github.com/njahn82/jct_data) and into hoaddata.
The JCT data used in the examined hoaddata version, which was enriched with linking ISSN (ISSN-L) and ROR-IDs from associated institutions before inclusion, cover agreements active from July 2021 to July 2024. -->

<!-- The JCT journal data was enriched with the linking ISSN (ISSN-L) according to the ISSN Registry Agency.
A comprehensive exclusion of fully open access journals was performed, following the methodology described in @jahn_2025.
This involved checking multiple journal lists (DOAJ, OpenAlex, Bielefeld) and article-level investigations.
The JCT institution data was enriched with ROR-IDs from associated institutions, such as university hospitals or institutes of large research organisations such as the Max Planck Society, according to OpenAlex' institution entity.
<!-- In total, xxxx hybrid journals could be linked to xxx agreements listed in the ESAC Transformative Agreements Registry, which formed to basis for obtaining article-level data including author affiliations, open access status ans document types. -->
The article metadata included DOIs, publication dates, open access information as well as author roles and institutional affiliations.
Publication years were determined using the earliest known date of publication in a journal.
In hoaddata, this corresponded to Crossref's issued date.<!--, although variations were observed (check ref)-->
For Web of Science and Scopus, the earliest publication date was used where available, with Scopus dates specifically determined by the KB through version tracking of the raw data.

Many transformative agreements typically cover only certain types of journal
articles, in particular original research articles including reviews.
Because of limited information on these document types in open scholarly data [@Haupka_2024], hoaddata used an extended version of Unpaywall's paratext recognition approach to exclude non-scholarly content [@Piwowar_2018].
To exclude conference supplements, which are also often not covered by transformative agreements, only articles published in regular issues, indicated by numerical pagination, were considered.
For Web of Science and Scopus, their established document type classifications were used to identify original research articles and reviews, referred to as original articles throughout this study.

@@

Articles in hybrid journals were considered open access when they were made freely available under a Creative Commons license on publishers' platforms. 
While hoaddata sourced this information from Crossref license metadata, Web of Science and Scopus relied on Unpaywall as evidence source. 
Unpaywall supplements Crossref metadata by parsing publisher websites directly, addressing cases where publishers do not provide machine-readable Creative Commons license information [@Piwowar_2018]. 
This additional parsing remains necessary despite transformative agreement workflows recommend the depositing license information during DOI registration [@Geschuhn_2017].
Both Web of Science and Scopus defined hybrid open access consistently as content available under Creative Commons licenses on publisher platforms, distinguishing it from bronze open access that lack such explicit license information, or use publisher-specific licenses.

@@

```{r cache=TRUE}
# Article level data TA
ta_oa_inst <- readr::read_csv(here::here("data-raw", "ta_oa_inst.csv"))
```
Author affiliations were retrieved for both first and, if available, corresponding authors to prepare the linking between articles and institutions covered by transformative agreements.
To handle different address variants, database-specific affiliation identifiers were used: ROR-IDs from OpenAlex for hoaddata, affiliation enhanced names for Web of Science, and Scopus Affiliation Identifiers for Scopus.
Additionally, ISO country codes were retrieved for each author's address to compile country-level statistics. 
These country codes for Web of Science and Scopus were provided by the KB.

Because neither Web of Science nor Scopus support ROR IDs, the institution identifier used by the JCT, a two-step matching process was implemented to harmonise affiliation data.
First, `r length(unique(ta_oa_inst$doi))` articles from `r length(unique(ta_oa_inst$ror))` institutions with ROR-IDs in the JCT data since 2017 (according to hoaddata) were processed to map first authors' ROR-IDs to corresponding proprietary affiliation identifier in Web of Science and Scopus using DOI matching.
Then, an algorithm selected the most frequent ROR ID and proprietary identifier pairs to handle multiple affiliations and organisational hierarchy differences.

@@

As a result of the comprehensive data processing described above, datasets on open access in hybrid journals included in transformative agreements were aggregated for each database at country and journal level by year.
@tbl-methods_overview_tab provides a general overview of the coverage between 2019 and 2023.
It shows that the majority of hybrid journals published at least one original research article or review marked as original during the five-year period.
These journals formed the basis for the subsequent calculation of article-level indicators.

```{r methods_overview_tab}
#| label: tbl-methods_overview_tab
#| tbl-cap: "Coverage of hybrid journals in transformative agreements 2019-23."
load(here::here("data", "manuscript_data.Rda"))
methods_overview_table_df <- bind_rows(
  # Journals
  jns_ind,
  # Articles
  articles_ind,
  # OA
  oa_articles_ind,
  # Affiliation and Auhtor roles
  aff_ind
) |>
  # Human-readable indicator labels
  mutate(
    indicator = case_when(
      indicator == "ind_active_jns" ~ "Active journals",
      indicator == "ind_active_jns_core" ~ "Active journals (core)",
      indicator == "ind_active_jns_core_with_oa" ~ "Active journals (core) with OA",
      indicator == "ind_articles_total" ~ "Total published articles ",
      indicator == "ind_articles_core" ~ "Original articles",
      indicator == "ind_articles_doi" ~ "Articles with DOI",
      indicator == "ind_articles_core_doi" ~ "Original articles with DOI",
      indicator == "ind_oa_articles_all" ~ "OA articles",
      indicator == "ind_oa_articles_core" ~ "Original OA articles",
      indicator == "first_aff" ~ "First author articles",
      indicator == "cor_aff" ~ "Corresponding author articles",
    )
  )

kableExtra::kbl(
  methods_overview_table_df,
  booktabs = TRUE, 
  format.args = list(big.mark = ","), 
  digits = 1, 
  escape = TRUE, 
  position = "H",
  col.names = c("", "HOAD", "Web of Science", "Scopus")
) |>
  kable_styling(position = "center") |>
  row_spec(0, bold = TRUE) |>
  pack_rows("Hybrid journal coverage", 1, 3) |>
  pack_rows("Publication volume", 4, 5) |>
  pack_rows("Digital Object Identifier (DOI) coverage", 6, 7) |>
  pack_rows("Open Access (OA) metrics", 8, 9) |>
  pack_rows("Original articles with affiliation data", 10, 11) 
```

While hoaddata only covers articles with a DOI, Scopus and Web of Science publication indicators were calculated using their database identifier.
A subsequent comparison of DOI coverage shows that non-original articles in Web of Science often lacked a DOI.
This was particularly the case for meeting abstracts, which are notably prevalent in Health Sciences journals [@Melero_Fuentes_2025] and are not indexed by Scopus [@Donner_2017].
Open access indicators were aggregated by DOI, as Unpaywall only collects information on open access status for articles with a DOI.
A closer look at original articles with affiliation data reveals a lack of data on corresponding authors in the case of OpenAlex, the affiliation data source used by hoaddata, compared to Web of Science and Scopus.
Only about two-thirds of the articles examined provided corresponding author affiliations.
For first authors, the proportion was `r methods_overview_table_df |> filter(indicator == "First author articles") |> pull(hoad) / methods_overview_table_df |> filter(indicator == "Core articles with DOI") |> pull(hoad) * 100`%.
Therefore, only first author data for hoaddata were considered in the following analysis.

@@

upset_overview <- ComplexUpset::upset(
  upset_data,
  c("Crossref", "WoS", "Scopus"),
  # Only journals in Crossref
  min_size = 30,
  set_sizes = FALSE,
  stripes = "transparent",
  # wrap = TRUE,
  name = "Data Source Coverage",
  base_annotations = list(
    "Journals" = (
      ComplexUpset::intersection_size(
        counts = FALSE,
        text = list(size = 8 / .pt),
        text_mapping = aes(
          label = paste0(!!upset_text_percentage(), '\n(', format(!!size, big.mark = ","), ')'),
          colour = ifelse(!!size > 5000, 'on_bar', 'on_background'),
          y = ifelse(!!size > 5000, !!size - 4000, !!size)
        ), width = bar_width
      ) +
        scale_y_continuous(labels = scales::number_format(big.mark = ",")) +
        ggtitle("A")
    )
  ),
  annotations = list(
    "Journal size\n(log scale)" = (
      ggplot(mapping = aes(x = intersection, y = overlap)) +
        geom_boxplot(outliers = FALSE, width = bar_width) +
        scale_y_log10(labels =  scales::number_format(big.mark = ","), limits = c(5, 10000)) +
        ggtitle("C")
    ),
    "Articles" = (
      ggplot(mapping = aes(y = n_cr)) +
        geom_bar(
          stat = "identity",
          color = "grey80",
          width = bar_width,
          aes(fill = "Surplus")
        ) +
        geom_bar(
          aes(y = overlap, fill = "Shared"),
          stat = "identity",
          width = bar_width,
          color = "#0091c8"
        ) +
        scale_fill_manual(
          values = c(
            `Surplus` = "grey80",
            `Shared` = "#0091c8"
          ),
          name = "DOI"
        ) +
        scale_y_continuous(labels =  scales::number_format(big.mark = ","),
                           breaks = c(0, 2500000, 5000000, 7500000)) +
        guides(fill = guide_legend(reverse = TRUE)) +
        ggtitle("B")
    )
  ),
  matrix = (intersection_matrix(geom = geom_point(
    shape = "square", size = 2
  )))
)
upset_overview & my_upset_theme
```
Journal coverage analysis revealed that `r round(nrow(union_jns) / nrow(all_jns) * 100)`% (n = `r nrow(union_jns)`) of hybrid journals included in transformative agreements were indexed in all three databases (Figure \ref{fig-upset_coverage_results}A).
The second-largest set consisted of journals indexed in both hoaddata and Scopus, comprising `r round( jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |>  nrow()`) of hybrid journals.
Notably, `r round(jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |>  nrow()`) of journals were exclusively contained in hoaadata, while another `r round( jn_upset_articles |> filter(Crossref == FALSE, Scopus == TRUE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(Crossref == FALSE, Scopus == TRUE) |>  nrow()`) were only found in the proprietory database Scopus. 
Of these, `r jn_upset_articles |> filter(WoS == TRUE, Crossref == FALSE, Scopus == TRUE) |>  nrow()` were also available in the Web of Science.
Upon inspection, this group of hybrid journals exclusively covered in the proprietary data sources mainly represented hybrid journals for which no open access evidence could be retrieved from Crossref, the open access evidence source for hoaddata.

In terms of article coverage, Figure \ref{fig-upset_coverage_results}B shows the total publication volume per combination in terms of DOI availability.
The largest set of hybrid journals, which includes all three data sources, also contains the largest number of articles. 
In total, these journals recorded `r upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum()` overlapping articles, represented by the blue bar. They represented `r round(upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum() / methods_overview_table_df |> filter(indicator == "Core articles with DOI") |> pull(wos) * 100)`% of original articles with DOI indexed in the Web of Science sample, and `r round(upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum() / methods_overview_table_df |> filter(indicator == "Core articles with DOI") |> pull(scp) * 100)`% in Scopus.
Another `r upset_data |> filter(Crossref == TRUE, WoS == FALSE, Scopus == TRUE) |> pull(n_cr) |> sum()` articles were exclusive to both Scopus and hoaddata. Exclusively in hoaddata were `r upset_data |> filter(Crossref == TRUE, WoS == FALSE, Scopus == FALSE) |> pull(n_cr) |> sum()` articles, and exclusively in the proprietary databases were `r upset_data |> filter(Crossref == FALSE, Scopus == TRUE) |> pull(overlap) |> sum(na.rm = TRUE)` articles.

@@

Figure \ref{fig-upset_coverage_results}B also shows the surplus of articles with DOI that were only available via hoaddata (grey area).
In case of hybrid journals covered by all three data sources, `r miss_dois |> nrow()` DOIs were only present in hoaddata.
After validation at the DOI level using the KB databases and manual inspection, the main reasons for missing DOI coverage in the proprietary database were insufficient classification of journal content as original research articles and reviews during the compilation of hoaddata. 
Particularly, letters and editorials could not be fully detected.
Paratext recognition failed in `r round(miss_dois_val |> filter(is.na(scp_item_type), is.na(wos_item_type)) |> pull(n_doi) / miss_dois |> nrow() * 100)`% of DOIs to identify non-scholarly content such as front matters or reviewer lists, which are generally not indexed by Scopus and Web of Science. 
To a lesser extent, differences in publication and indexing dates were a reason for non-overlapping DOIs.

Using overlapping DOIs, the publication volume between 2019 and 2023 was also calculated for each journal. 
Figure \ref{fig-upset_coverage_results}C illustrates the distribution for each combination. It shows that there is a large spread across the journals covered by all three data sources. 
Furthermore, journals in this set published more on average than journals covered by only one or two data sources. 
In particular, the journals covered exclusively by hoaddata were substantially smaller than those covered by all three sources. 
Upon inspection, these were often newly launched journals, which explains the relatively low five-year publication volume.
An example is *Digital society* that published `r jn_upset_articles |> filter(issn_l == "2731-4650") |> pull(n_cr)` articles.
This hybrid journal was launched in 2022, being covered by various Springer Nature transformative agreements since then.

@@

```{r oa_country_comparision_plot_, fig.asp=1, fig.width = 8}
#| fig-cap: "Open Access by country"
#| label: fig-uptake_country
#| 
layout <- "EE
AC
BD"
